Classification of Parkinson's disease patients using Bayesian deep learning based on fMRI data

### 2020-02-22 article.tex
citations improved

Parkinson's disease is a long-term degenerative disorder of the central nervous system that mainly affects the motor system. As the disease worsens, non-motor symptoms become more common. The symptoms usually emerge slowly. Early in the disease, the most obvious symptoms are shaking, rigidity, slowness of movement, and difficulty with walking. Thinking and behavioral problems may also occur. Recently, resting-state fMRI has become an especially popular area of research, and, as a result, the publication of related articles has grown exponentially \cite{Murphy2013}. The goal of resting-state fMRI is to use the common variance of the fMRI blood oxygenation level dependent (BOLD) signals in different regions of the brain as an indicator of synchronous neural activity. The assumption is that the temporal similarity between the BOLD signals in each region demonstrates that they are in constant communication with one another and thus form a functional network \cite{Calhoun2016}.

There are many ways to define network nodes from fMRI; nodes are often defined as spatial regions of interest, for example, as obtained from fMRI activation or from brain atlases. Alternatively, parcellation via a data-driven clustering of the FMRI data itself (e.g., hierarchical clustering or independent component analysis) can be run to define clusters or components (spatial maps with associated time-slices), which can be considered network nodes \cite{Smith2012}. Furthermore, we are curious to find out if fMRI scans carry enough information without taking their temporal connectivities into account. In other words, does spatial information carry enough data to diagnose Parkinson's disease patients?

@@

Take medical diagnosis as an example. Besides seeing
visible symptoms (or medical images from MRI) and hearing descriptions from patients, a doctor has to look for relations among all the symptoms and preferably infer the corresponding etiology. Only after that can the doctor provide medical advice for the patients. In this example, although the abilities of seeing and hearing allow the doctor to acquire information from the patients, it is the thinking part that defines a doctor. Specifically, the ability of thinking here could involve causal inference, logic deduction, and dealing with uncertainty, which is apparently beyond the capability of conventional deep learning methods \cite{Wang2016}.

@@

In this study, we showed that spatial information is powerful enough to diagnose Parkinson's disease. Moreover, we can try to implement a hybrid solution based on the proposed method that uses temporal connectivities as well. For instance, we can use functional connectivity based algorithms \cite{Rashid2016} \cite{Shen2014} \cite{Calhoun2004} \cite{GonzalezCastillo2015} \cite{Greicius2002} \cite{Arribas2010} \cite{Leonardi2013} \cite{Liu2015} \cite{Liu2016} \cite{Lehmann2017} \cite{Ryali2012} or statistical/Bayesian based methods \cite{HanchuanPeng2005} \cite{Flandin2007} \cite{Ahmad2016} \cite{Power2012} \cite{Oikonomou2010} \cite{GomezLaberge2011} \cite{Zhang2001} \cite{Oikonomou2012} \cite{Oikonomou2013} \cite{Younes2006} \cite{Suk2016} \cite{Calhoun2012}. Ensembling can be done by training mentioned methods combined with Bayesian units or by adding their extracted features to the existing network.

>>>>

Parkinson's disease is a long-term degenerative disorder of the central nervous system that mainly affects the motor system. As the disease worsens, non-motor symptoms become more common. The symptoms usually emerge slowly. Early in the disease, the most obvious symptoms are shaking, rigidity, slowness of movement, and difficulty with walking. Thinking and behavioral problems may also occur. Recently, resting-state fMRI has become an especially popular area of research, and, as a result, the publication of related articles has grown exponentially \cite{Murphy2013}. The goal of resting-state fMRI is to use the common variance of the fMRI blood oxygenation level dependent (BOLD) signals in different regions of the brain as an indicator of synchronous neural activity \cite{GomezLaberge2011}\cite{Oikonomou2010}. The assumption is that the temporal similarity between the BOLD signals in each region demonstrates that they are in constant communication with one another and thus form a functional network \cite{Calhoun2016}.

There are many ways to define network nodes from fMRI; nodes are often defined as spatial regions of interest, for example, as obtained from fMRI activation or from brain atlases \cite{Arribas2010}\cite{Calhoun2012}. Alternatively, parcellation via a data-driven clustering of the FMRI data itself (e.g., hierarchical clustering or independent component analysis) can be run to define clusters or components (spatial maps with associated time-slices), which can be considered network nodes \cite{Smith2012}. Furthermore, we are curious to find out if fMRI scans carry enough information without taking their temporal connectivities into account. In other words, does spatial information carry enough data to diagnose Parkinson's disease patients?

@@

Take medical diagnosis as an example. Besides seeing visible symptoms (or medical images from MRI) and hearing descriptions from patients, a doctor has to look for relations among all the symptoms and preferably infer the corresponding etiology. Only after that can the doctor provide medical advice for the patients \cite{GonzalezCastillo2015}. In this example, although the abilities of seeing and hearing allow the doctor to acquire information from the patients, it is the thinking part that defines a doctor. Specifically, the ability of thinking here could involve causal inference, logic deduction, and dealing with uncertainty, which is apparently beyond the capability of conventional deep learning methods \cite{Wang2016}.

@@

In this study, we showed that spatial information is powerful enough to diagnose Parkinson's disease. Moreover, we can try to implement a hybrid solution based on the proposed method that uses temporal connectivities as well. For instance, we can use functional connectivity based algorithms \cite{Lehmann2017}\cite{Leonardi2013}\cite{Liu2016}\cite{Liu2015}\cite{Ryali2012} or statistical/Bayesian based methods \cite{Ahmad2016}\cite{Oikonomou2013}\cite{Oikonomou2012}\cite{Power2012}\cite{Suk2016}. Ensembling can be done by training mentioned methods combined with Bayesian units or by adding their extracted features to the existing network.
